---
title: "Exploratory Analysis Showcase"
---

This page provides a quick showcase of exploratory analysis using the cleaned SSI-2012 dataset. We illustrate basic data distributions and aesthetic tastes using clean, standard visualizations in R.

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(haven)

# Load the cleaned dataset
df <- readRDS("../data/clean/ssi2012_cleaned.rds")
```

## 1. Classical Music Tastes

Let's examine how respondents rate their liking of **Classical music** (`classicaltaste`). This variable has four main levels: *like*, *dislike*, *neither like nor dislike*, and *I am not familiar with this type of music*.

```{r}
#| label: fig-classical
#| fig-cap: "Distribution of Classical Music Tastes among respondents."
#| warning: false
#| message: false

# Convert classicaltaste to standard factor for plotting
df_classical <- df |> 
  filter(!is.na(classicaltaste)) |> 
  mutate(classicaltaste = as_factor(classicaltaste))

# Create bar chart with aesthetic mapping to y-axis (avoiding coord_flip)
ggplot(df_classical, aes(y = classicaltaste)) +
  geom_bar() +
  labs(
    x = "Count of Respondents",
    y = "Classical Music Taste"
  )
```

As shown in @fig-classical, a substantial majority of the 2,276 respondents report that they like Classical music, while only a small percentage are completely unfamiliar with it.

---

## 2. Subjective Social Class Self-Identification

Respondents were asked: *"Which of these do you think best describes you?"* with options: *lower class*, *working class*, *middle class*, and *upper class* (`percclass`).

```{r}
#| label: fig-class
#| fig-cap: "Self-described subjective social class of respondents."
#| warning: false
#| message: false

# Convert percclass to factor and filter missing values
df_class <- df |> 
  filter(!is.na(percclass)) |> 
  mutate(percclass = as_factor(percclass))

# Plot distribution
ggplot(df_class, aes(y = percclass)) +
  geom_bar() +
  labs(
    x = "Count of Respondents",
    y = "Subjective Social Class"
  )
```

As displayed in @fig-class, the majority of respondents identify as middle class or working class, with smaller proportions self-identifying as lower class or upper class.

---

## 3. Painting Impression Ratings (Pre-Test)

Before receiving any social manipulation, respondents gave their first impression of an abstract painting on a scale from 1 (dislike very much) to 7 (like very much) (`taste1`).

```{r}
#| label: fig-painting-taste
#| fig-cap: "Distribution of initial painting taste ratings (1-7 scale)."
#| warning: false
#| message: false

# Convert taste1 to standard factor to represent its ordinal categories
df_painting <- df |> 
  filter(!is.na(taste1)) |> 
  mutate(taste1_factor = as_factor(taste1))

ggplot(df_painting, aes(y = taste1_factor)) +
  geom_bar() +
  labs(
    x = "Count of Respondents",
    y = "Initial Rating (1 = Dislike Very Much, 7 = Like Very Much)"
  )
```

The pre-test ratings shown in @fig-painting-taste reveal a wide distribution of opinions on the painting, peaking in the neutral and mild liking ranges.
